---
title: "Tugas Kelompok 2"
author: "Kelompok S"
date: "2023-09-26"
output:
  html_document:
    df_print: paged
  pdf_document: default
---
Nama Kelompok :

-   Andreas Hendra Herwanto (164221064)

-   Fabyan Riza Kiram (164221068)

-   Octaviani Putri Al Fajri (164221042)


## Data Source

<https://www.kaggle.com/datasets/ibrahimelsayed182/plane-ticket-price>


## Data Description

1 . ID: Contiguous sample number

2 . Airline: The name of the airline

3 . DateofJourney: The date of the journey

4 . Source: The source from which the service begins

5 . Destination: The destination where the service ends

6 . Dep_Time: The time when the journey starts from the source.

7 . Arrival_Time: Time of arrival at the destination.

8 . Duration: Total duration of the flight.

9 . Total_Stops: Total stops between the source and destination.

10 . Additional_Info: Additional information about the flight

11 . Price: The price of the ticket

## Import Data

```{r}
setwd("D:/KULIAH/Tugas/SEMESTER 3/EVD")

library(readr)

df=read_csv("plane.csv")
head(df)
```

## Variabel Selection
```{r}
library(dplyr)

df_clear<-select(df,-c("Date_of_Journey","Route","Dep_Time","Arrival_Time","Duration","Additional_Info"))
head(df_clear)
```

## Checking Missing Value
```{r}
missing_values_per_column <- colSums(is.na(df_clear))
missing_values_per_column

```

## Label Encoding
```{r}
library(CatEncoders)

df_encoding<-data.frame(df_clear)

labs1 = LabelEncoder.fit(df_encoding$Airline)
labs2 = LabelEncoder.fit(df_encoding$Source)
labs3 = LabelEncoder.fit(df_encoding$Destination)

df_encoding$Airline <- transform(labs1, df_encoding$Airline)
df_encoding$Source <- transform(labs2, df_encoding$Source)
df_encoding$Destination <- transform(labs3, df_encoding$Destination)

df_encoding
```

## Correlation Table

```{r}
library(corrplot)
corrplot(cor(df_encoding),
         order = "hclust",
         tl.col="black", 
         tl.srt=45, 
         method = "number") 
```
**Keterangan :**

-   Nilai korelasi negatif : Kedua variabel memiliki hubungan yang berlawanan.

-   Nilai korelasi positif : Kedua variabel memiliki hubungan yang searah.

-   Nilai korelasi antara 0 -- 0,25 : Korelasi sangat lemah

-   Nilai korelasi antara 0,25 -- 0,5 : Korelasi cukup

-   Nilai korelasi antara 0,5 -- 0,75 : Korelasi kuat

-   Nilai korelasi antara 0,75 -- 0,99 : Korelasi sangat kuat

-   Nilai korelasi antara 1 atau -1 : Korelasi hubungan sempurna

## EDA

```{r}
library(dplyr)
library(ggplot2)
library(ggrepel)
```

1. Proporsi Jumlah Tiket dari Setiap Maskapai

```{r}
plotdata<-df %>%
  count(Airline) %>%
  arrange(desc(Airline)) %>%
  mutate(prop = round(n*100/sum(n),1),
         lab.ypos=cumsum(prop)-0.5*prop)

ggplot(plotdata,aes(x="",y=prop, fill=Airline))+
  geom_bar(width=1,
           stat="identity",
           color="black",
           alpha=0.8)+
  coord_polar('y', start=0,
              direction=-1)+
  theme_void()+
  labs(title= "Proporsi Jumlah Tiket dari Setiap Maskapai")+
  theme(plot.title = element_text(hjust = 0.8, vjust=-5))+
  theme_void()+
  scale_fill_brewer(palette = "Set3") 
  
```
**Interpretasi :**

2. Source dan Destinasi Penerbangan di India selama 2019
```{r}
ggplot(df, aes(y=Source, fill=Destination))+
  geom_bar()+
  labs(title="Penerbangan Maskapai India Selama 2019")+
  scale_fill_brewer(palette = "Set3")

```

3. Density Variabel Dependen (Price(Rupee))
```{r}
a<-ggplot(df, aes(`Price(Rupee)`))+
  geom_histogram(aes(y = ..density..),
                 fill = "#EEDD82",
                 alpha=0.8)+
  geom_density(color='black',lwd=0.5)+
  labs(title="Distribusi Biaya Asuransi (Charges)")+
  theme(plot.title = element_text(hjust = 0.8))

b<-ggplot (df, aes(y=`Price(Rupee)`)) +
  geom_boxplot (fill = "#EEDD82",
                alpha = 0.6,
                outlier.size=2,
                varwidth=TRUE) +
  labs (title = "Charges distribution")


library(gridExtra)
grid.arrange(a, b, ncol = 2)
```
**Interpretasi :**


4. Trend Harga Tiket Maskapai di India selama 2019

```{r}

data <- data.frame(
  Date = as.Date(df$Date_of_Journey, "%d/%m/%Y"),
  Price = df$`Price(Rupee)`)

test = aggregate(`Price(Rupee)`~Date_of_Journey,df,mean)

# Create a line chart with multiple lines (one line per flight)
ggplot(test, aes(x = as.Date(Date_of_Journey, "%d/%m/%Y"), y = `Price(Rupee)`))+
  geom_line(lwd=0.6, color="#EE2C2C") +
  labs(
    title = "Line Chart of Price Over Time for Multiple Flights",
    x = "Date",
    y = "Price")

```
**Interpretasi :**


5.  Plot Price vs Total Stops and Price vs Duration (Hour)

```{r}
plotdata5 <- df%>%
  group_by(Total_Stops)%>%
  summarize(max=max(`Price(Rupee)`))

a1<-ggplot(plotdata5, aes(x=Total_Stops, y=max))+
  geom_bar(stat="identity", fill="lightskyblue")+
  labs(title="Price(Rupee) vs Total Stops",
       y= "Price(Rupee)")

b1<-ggplot(df, aes(x=`Duration(Hour)`, y=`Price(Rupee)`))+
  geom_point(color="#FFA07A", size=1.3)+
  labs(title="Price(Rupee) vs Duration (Hour)")

library(gridExtra)
grid.arrange(a1, b1, ncol = 2)
```

6. Perbadingan Harga Setiap Maskapai
```{r}
ggplot(df,aes(x=`Price(Rupee)`,y=Airline))+
  geom_boxplot(alpha=.7,  outlier.size=1, color="#CD2626")+
  labs(title = "Price distribution by Airline")


```


